---
title: "P8105 Data Science I Homework 6"
author: Olivia Wang (hw2852)
output: github_document
date: "2022-12-03"
---

In preparation for the problems below, we will load the following libraries: 

```{r load_libraries}
library(tidyverse)
library(readxl)
library(dplyr)
```

# Problem 1



# Problem 2

## 2.1 Homicide Data Import and Cleaning

Let us begin by importing the CSV file containing _Washington Post's_ homicide data downloaded from GitHub, and applying the `clean_names` function. Next, we can create a new `city_state` variable (e.g. "Baltimore, MD") by joining the existing city and state variables using the `paste` command, and a new binary `homicide_solved` variable indicating whether a homicide is solved. The `victim_age` variable previously read as a character variable is transformed into a numeric variable. Finally, we filtered the data to exclude information from Dallas, TX, Phoenix, AZ, Kansas City, MD, and Tulsa, AL, and to only include victims whose reported race was White or Black. 

```{r}
homicide_data = 
  read_csv("./homicide-data.csv") %>% 
  janitor::clean_names() %>% 
  mutate(city_state = as.character(paste(city, state, sep = ", ")), 
         homicide_solved = ifelse(disposition == "Closed by arrest", 1, 0), 
         victim_age = as.numeric(victim_age)) %>% 
  filter(!(city_state %in% c("Dallas, TX", "Phoeniz, AZ", "Kansas City, MO", "Tulsa, AL")), 
         victim_race %in% c("White", "Black"))
```

## 2.2 Homicide Data Analysis: Odds Ratios & 95% CIs

#### _Solved Homicides Comparing Male & Female Victims in Baltimore, MD_

Below we use the `glm` function to estimate the odds ratio of solved homicides comparing male to female victims, specifically in Baltimore, MD, adjusting for age and race. The arguments for the `glm` function indicate that the `homicide_solved` variable is the outcome of interest, and `victim_age`, `victim_sex` and `victim_race` are the predictors.

The output generated from the `glm` function is then saved as an R object. The `broom::tidy()` function was applied to tidy the output, and the log odds ratios, odds ratios, and p-value corresponding to each term in the regression model were pulled and tabulated. In the final step, the `confint` function was applied to the output generated from the `glm` function to generate the 95% confidence intervals surrounding the odds ratios. 

```{r}
homicide_data_df = homicide_data %>% 
  filter(city_state == "Baltimore, MD")

glm_homicide_baltimore = 
  homicide_data_df %>% 
  glm(homicide_solved ~ victim_age + victim_sex + victim_race, data = ., family = binomial())

glm_homicide_baltimore %>%
  broom::tidy() %>%
  mutate(odds_ratio = exp(estimate), 
         lower_CI = exp(estimate - 1.96*std.error), 
         upper_CI = exp(estimate + 1.96*std.error)) %>%
  select(term, odds_ratio, lower_CI, upper_CI) %>% 
  filter(term == "victim_sexMale") %>% 
  knitr::kable(col.names = c('Term', 'Odds Ratio', 'Lower 95% CI Limit', 'Upper 95% CI Limit'), digits = 3)
```

In Baltimore, MD, the odds ratio of solving homicides comparing male victims to female victims, adjusting for age and race, was 0.456. We are 95% confident that the true odds ratio of solving homicides comparing male and female victims, adjusting for age and race, lies between 0.340 and 0.612.

#### _Solved Homicides Comparing Male & Female Victims in All Cities_

We will now apply the `glm` function to all cities in the data set. This process involves first creating a new `glm_homicide` function, which applies the `glm` function to selected inputs. We then use the `select` function to reorder the columns in the data frame to have `city_state` appear first, as the function will be mapped to each unique `city_state`. The city-level homicide data will then be nested using the `nest` function to generate list columns for city-level homicide counts, and the `glm_homicide` function will be mapped to each tibble using `purrr::map`. Finally, we apply the `unnest` function to generate the city-specific odds ratio and 95% CIs. 

```{r}
glm_homicide = function(homicide_data) {
  glm(homicide_solved ~ victim_age + victim_sex + victim_race, data = homicide_data, family = binomial()) %>% 
  broom::tidy() %>%
  mutate(odds_ratio = exp(estimate), 
         lower_CI = exp(estimate - 1.96*std.error), 
         upper_CI = exp(estimate + 1.96*std.error)) %>%
  select(term, odds_ratio, lower_CI, upper_CI) %>% 
  filter(term == "victim_sexMale")
}

homicide_data_analysis = homicide_data %>% 
  select(city_state, everything()) %>% 
  nest(data = uid:homicide_solved) %>% 
  mutate(glm_homicide_output = purrr::map(.x = data, ~ glm_homicide(.x))) %>% 
  unnest(cols = glm_homicide_output)
```

#### _Plotting City-Specific Solved Homicides Comparing Male & Female Victims_

using the generated output from the analysis above, we will plot the city-level odds ratio estimates of unsolved homicides comparing male to female victims and corresponding 95% CIs using `ggplot`. In addition to plotting the OR estimates using `geom_point`, the 95% confidence intervals associated with each estimate were applied using `geom_errorbar`. Cities are ordered in increasing OR estimates of unsolved homicides, comparing male to female victims. 

```{r}
homicide_data_analysis %>% 
  ggplot(aes(x = reorder(city_state, odds_ratio), y = odds_ratio, color = city_state)) + 
  geom_point() +
  geom_errorbar(aes(ymin = lower_CI, ymax = upper_CI)) +
  labs(
    title = "City-Level Odds Ratio Estimates and 95% CIs for Unsolved Homicides Comparing Male to Female Victims", 
    x = "City, State", 
    y = "Odds Ratio Estimate") + 
  theme(
    axis.text.x = element_text(angle = 70, hjust = 1), 
    legend.position = "none")
```

# Problem 3 

## 3.1 Birth Weight Data Import and Cleaning

Let us begin by importing the CSV file containing birth weight data, and applying the `clean_names` function. The following variables were converted from numeric to factor variables to reflect the information in the codebook: `babysex`, `frace`, `malform`, and `mrace`. The `fincome` variable, denoting family monthly income, was also modified to reflect the actual reported amounts (i.e., hundreds). Finally, the `skim` function was applied to generate a summary of the variables in the data set. 

```{r}
birthweight_data = 
  read_csv("./birthweight.csv") %>% 
  janitor::clean_names() %>% 
  mutate(babysex = factor(babysex, labels = c("Male", "Female")),
         fincome = fincome * 100, 
         frace = factor(frace, labels = c("White", "Black", "Asian", "Puerto Rican", "Other")),
         malform = factor(malform, labels = c("Absent", "Present")),
         mrace = factor(mrace, labels = c("White", "Black", "Asian", "Puerto Rican")))

skimr::skim(birthweight_data)
```

Based on the results from applying the `skim` function, there are no missing values for any variable in the data set. 

## 3.2 Birth Weight Data Analysis: Modeling Exercise



